---
title: "Accessing APIs"
author: "Uni Lee"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(httr)
library(rtweet)
library(jsonlite)
library(rmarkdown)
```

## APIs

*Application Programming Interface (API)* is a defined interface for communication between software components.
- *Web* API: provides an interface to structured data from a web service
- APIs is an easier way to collect mass data from the web than webscraping. So, you're lucky if your data provider has APIs! 

## Functionality of APIs

Consider the API as the server, and yourself as the client. You can communicate with the server using API endpoints (URLs).

*Hypertext transfer protocol (HTTP)* is the APIs' method of communication, which contains header and message body. 

Here are some messages that you'd get:

- 2xx: Success
- 4xx: Client error (incl. the popular *404: Not found* or *403: Forbidden*)
- 5xx: Server error
- The message body contains the requested data in a specific format, often JSON or XML.

## Examples of popular APIs

Social media:

- [Twitter](https://developer.twitter.com/)
- [Facebook Graph API](https://developers.facebook.com/docs/graph-api/) (restricted to own account and public pages)
- [YouTube (Google)](https://developers.google.com/youtube/)
- [LinkedIn](https://www.linkedin.com/developers/)
- [National Renewable Energy Laboratory](https://developer.nrel.gov/)
- For more, see [programmableweb.com](http://www.programmableweb.com/).

### Exercise: Getting utility rates from OpenEI
As an exercise, we will get data on the US utility rates from OpenEI API. You need the `httr` and `jsonlite` package for this.

First, identify base url and endpoints. Sign up to get a token for access (api_key). 

```{r}
base <- "https://api.openei.org/"
endpoint <- "utility_rates"
key <- "&api_key=0uDDzDsgxHMfZbIfqEbGHEpPqnCgNwWkEYIMj4KB"

query <- "?version=latest&format=json"

query_full <- paste0(base, endpoint,query, key)

res <- GET(query_full) # Ask the server for data 
res

```

The query was successful (status=200)! The data taks on a `json` format, so we will use functions from the `jsonlit`e library to get the results in a dataframe. 

```{r}
utilityrate <- rawToChar(res$content) %>% # Data in a character format (JSON structure) 
  fromJSON() # This returns a list

utilityrate_df <- utilityrate$items

names(utilityrate_df)

utilityrate_sub <- utilityrate_df %>% 
  select(sector, utility, name, source, country, startdate, description)
paged_table(utilityrate_sub)
```


3. Collect some data from their API from a politician running in your constituency (or a constituency of your choice). Tip: Use the `fromJSON` function in `jsonlite` to load a JSON file via the API into R. Convert the output into a nicely formatted dataframe.

## API wrapper packages

Use "API wrapper packages" in R whenever possible, because they are easier to use: *rtweet*, *ggmap* (geocoding via Google Maps), *wikipediR*, etc.

## Twitter

* REST APIs are used to read and write tweets. 
* Streaming APIs provides low latency access to 1% of global stream - public, user and site streams
- See documentation athttps://dev.twitter.com/overview/documentation

### Wrapper package "rtweet"
Find out what you can do with the `rtweet` package: <https://github.com/mkearney/rtweet> 

In general, the function `search_tweets` will be your friend. This function allows you to collect data from the past 6-9 days. 

For the purpose of this exercise, I will collect tweets with terms "plant-based meat" or "alternative protein". The first argument of `search_tweets` function is `q`, which takes queries as strings to search in Twitter's REST API.

```{r}
tweets <- search_tweets(
  "plant-based meat OR alternative protein", # Query to be searched
  n=1000, include_rts=FALSE # Only get original tweets 
)

length(names(tweets)) #The dataset provided by Tweeter API has 90 variables! 
tweets_sub <- tweets %>% select(text, user_id, name,hashtags, created_at)  #Select relevant variables
paged_table(tweets_sub)

```

